Collective behavior of Suspended Janus Particles via an Integral Equation Method

### 2022-05-12 doc/PRF2022/PRF_main.tex
updates

Suppose we use the double layer potential to express the solution of \eqref{eq:SL} where the condition is $f(u)=u^2$. Then it is given by 
\begin{align}
u({\bf x}) = \DD[\sigma](\xx) = \int_\Gamma \frac{\partial G(\xx-\yy)}{\partial \nnu_\yy}\sigma(\yy)\, \dif s_\yy,
\end{align}
%
where $G(\xx)$ is the fundamental solution to the screened Laplace equation \eqref{eq:SL}.
Since the free energy functional can be written as 
\begin{align}
\label{eq:free_energy2}
  E[u(\xx)] &= C\rho \int_\Gamma u \frac{\partial u}{\partial \nnu}  \,\dif s, \quad \xx\in\Gamma_i,
\end{align}
%
we can then use the following identity to calculate this boundary equation without evaluating 
the normal derivative of the double layer potential directly.

\begin{align}
\nnu_\xx \cdot \nabla_\zz u(\zz) &=\nnu_\xx \cdot \nabla_\zz \int_\Gamma \frac{\partial G(\zz-\yy)}{\partial \nnu_\yy}\sigma(\yy) \,\dif s_\yy\\
&=\int_\Gamma \nnu_\xx\cdot \left(\nabla_\zz\nabla_\yy^\top  G(\zz-\yy)\right)\cdot \nnu_\yy\sigma(\yy)  \,\dif s_\yy\\
&=-\int_\Gamma \nnu_\xx\cdot \left(\nabla_\yy\nabla_\yy^\top G(\zz-\yy)\right)\cdot\nnu_\yy\sigma(\yy)\ \dif s_\yy\\
&=-\int_\Gamma {\bf t}_\xx\cdot\Delta_\yy G(\zz-\yy) {\bf t}_\yy \sigma(\yy)\ \dif s_\yy+\int_\Gamma({\bf t}_\xx\cdot\nabla_\yy)({\bf t}_\yy\cdot\nabla_\yy G(\zz-\yy))\sigma(\yy)\ \dif s_\yy\\
&= -\int_\Gamma {\bf t}_\xx\cdot\frac1\rho^2 G(\zz-\yy) {\bf t}_\yy \sigma(\yy)\ \dif s_\yy - 
({\bf t}_\xx\cdot\nabla_\zz)\int_\Gamma \frac{\partial G(\zz-\yy)}{\partial s_\yy}\sigma(\yy)\ \dif s_\yy\\
&= -\frac1\rho^2 {\bf t}_\xx\cdot \int_\Gamma G(\zz-\yy){\bf t}_\yy \sigma(\yy)\ \dif s_\yy + 
({\bf t}_\xx \cdot \nabla_\zz)\int_\Gamma G(\zz-\yy)\frac{\partial \sigma(\yy)}{\partial s}\ \dif s_\yy.
\end{align}
%
Here ${\bf t}_\xx$ is the tangent vector at the point $\xx$ and $d/ds$ is the arclength derivative.
Letting $\zz\to\xx\in\Gamma$, we obtain
%
\begin{equation}
\frac{\partial}{\partial \nnu_\xx} \DD[\sigma](\xx) = -\frac1\rho^2 {\bf t}_\xx\cdot \SSS[\sigma{\bf t}](\xx) + \frac{\partial}{\partial s}\SSS\left[\frac{\partial \sigma}{\partial s}\right](\xx).
\end{equation}
%

@@

Distinct morphologies can be obtained from 
the HAP model by shifting the boundary condition $g(\xx)$.
Consider the linear case where $f(u) = u^2$.
This choice makes \eqref{eq:SL} a boundary value
problem for the screened-Laplace equation
and the solutions have a boundary layer structure that decays to $0$ in the
bulk with the decay length $\rho$.
We consider three kinds of boundary conditions
on particle $\Gamma_i$:
\begin{equation}
  \label{eq:bcs}
  \begin{tabular}{c|c|c}
     \text{(a)} & \text{(b)} & \text{(c)} \\
    \hline
    \rule{0pt}{4ex} 
      $g(\mathbf{x}) = \displaystyle \frac{1+\cos (\theta_i(\xx))}{\sqrt{3\pi r}}$
    & $g(\mathbf{x}) = \displaystyle \frac{2+\cos (\theta_i(\xx))}{\sqrt{9\pi r}}$
    & $g(\mathbf{x}) = \displaystyle \frac{\cos (\theta_i(\xx))}{\sqrt{\pi r}}$
\end{tabular}
\end{equation}
  The number $\theta_i(\xx)$ is the angle formed by the position $\xx$, the
particle center $\aa_i$, and the particle director $\dd_i$.% (Figure \ref{fig:flow_map}).
The normalizations e.g., $(\pi r)^{-1/2}$, provide a uniform
surface energy $\int_{\Gamma_i} g^2 \,ds = 1$ for circular particles
of radius $r$. 
The cases (b) and (c) are obtained by applying a vertical shift
and scaling to case (a), see Figure \ref{fig:bcs}.
Using this setup, we simulate the dynamics of 60 particles with random
initial position and orientation as shown in Figure~\ref{fig:relax}.

@@

\begin{figure}
  \begin{center}
  \includegraphics[width=0.4\textwidth]{Fig2a.png}
  \includegraphics[width=0.4\textwidth]{Fig2b.png}\\
  \includegraphics[width=0.4\textwidth]{Fig2c.png}
  \includegraphics[width=0.4\textwidth]{Fig2d.png}
  \end{center}
  \vspace{-20pt}  
  \caption{\label{fig:relax} (a) Initial configuration; (b)--(d) Equilibruim states for all boundary conditions~\eqref{eq:bcs}(a)--(c), respectively. Colors represent the boundary values that increases from minimum (blue) to maximum (red).  }
\end{figure}

Case (a) models amphiphilic particles.
Amphiphilic particles have a hydrophobic tail and a hydrophilic head
and these are accounted for as follows.
The hydrophilic side takes the value $u =0$.
This mimicks the apolar head of a lipid, for example, which does
not alter the structure of adjacent waters. The hydrophobic side
represents hydrocarbons and takes the value 
$u > 0$. The interaction between particles is attractive,
and particles will collectively orient their tails toward
one another. Figure~\ref{fig:relax}(b) demonstrates multiple bilayer structures such as pancake and vesicle shapes.

Multilamelar bilayers arise when both sides of the particle 
are hydrophobic. 
The boundary condition \eqref{eq:bcs}(b) 
gives a particle with a hydrophobic intensity that is greater
on the $\theta_i = 0$ side than on the $\theta_i = \pi$ side.
The initial self-assembly is similar to that in case (a).
The difference arises in the long-time dynamics where the bilayers
no longer remain well-separated. 
Rather, the bilayers form layers 
as a consequence of the interfacial tension of exposed particle
heads.%~\cite{Huetal19, deMeetal21}. 
Figure~\ref{fig:relax}(c) shows the multilamellar structure with 8 layers when $N_b=60$. 
The number of folds depends on a number of factors, for instances, number of particles and intial configurations.

Finally, boundary condition \eqref{eq:bcs}(c) models 
a particle whose head surface repels the tail surface as proposed in
\cite{MaRa76, Ma77}.
The particles initially form chains with their directors perpendicular to the
length of the chain. The equilibrium structure
resembles a checkerboard pattern
where each particle coordinates
its head with the head of three other particles and its tail with the
tail of three other particles. % Figure~\ref{fig:self-assembly}(c).
Figure~\ref{fig:relax}(d) gives an example that this boundary condition results a checker board/stripes equilibrium state.

@@

\begin{figure}
  \begin{center}
  \includegraphics[width=0.4\textwidth]{shear_0.png}
  \includegraphics[width=0.4\textwidth]{shear_1000.png}\\
  \includegraphics[width=0.4\textwidth]{shear_2000.png}
    \includegraphics[width=0.4\textwidth]{shear_dist.jpg}
  \end{center}
  \vspace{-20pt}  
  \caption{\label{fig:shear_1} Snapshots of simulation using boundary condition \eqref{eq:bcs}(b) in a shear flow when $\dot\gamma=0.05$. }
\end{figure}

@@

Consider placing equilibrium states of cases (b) and (c) for 60 particles in a shear flow,
\begin{equation}
\uu_\infty(\xx) = \dot\gamma ({\bf e}_y\cdot\xx){\bf e}_x,
\end{equation}
%
where $\dot\gamma$ is the shear rate.

We measure the end-to-end distance of the multi-lamellar structure when the flow is on.

>>>>

We adopt the double layer potential to express the solution of \eqref{eq:SL} where the condition is $f(u)=u^2$. Then it is given by 
\begin{align}
\label{eq:energy_BIE}
u({\bf x}) = \DD[\sigma](\xx) = \int_\Gamma \frac{\partial G(\xx-\yy)}{\partial \nnu_\yy}\sigma(\yy)\, \dif s_\yy,
\end{align}
%
where $G(\xx)$ is the fundamental solution to the screened Laplace equation \eqref{eq:SL} and $\nnu$ is the outward normal.
Since the free energy functional can be written as 
\begin{align}
\label{eq:free_energy2}
  E[u(\xx)] &= C\rho \int_\Gamma u \frac{\partial u}{\partial \nnu}  \,\dif s, \quad \xx\in\Gamma_i,
\end{align}
%
we can then refer to the identity (\cite{Hsiao2008}, \S 1.2) to calculate this boundary integral equation without evaluating the normal derivative of the double layer potential directly.

Consider
\begin{align}
\nnu_\xx \cdot \nabla_\zz u(\zz) &=\nnu_\xx \cdot \nabla_\zz \int_\Gamma \frac{\partial G(\zz-\yy)}{\partial \nnu_\yy}\sigma(\yy) \,\dif s_\yy\\
&=\int_\Gamma \nnu_\xx\cdot \left(\nabla_\zz\nabla_\yy^\top  G(\zz-\yy)\right)\cdot \nnu_\yy\sigma(\yy)  \,\dif s_\yy\\
&=-\int_\Gamma \nnu_\xx\cdot \left(\nabla_\yy\nabla_\yy^\top G(\zz-\yy)\right)\cdot\nnu_\yy\sigma(\yy)\ \dif s_\yy\\
&=-\int_\Gamma {\bf t}_\xx\cdot\Delta_\yy G(\zz-\yy) {\bf t}_\yy \sigma(\yy)\ \dif s_\yy+\int_\Gamma({\bf t}_\xx\cdot\nabla_\yy)({\bf t}_\yy\cdot\nabla_\yy G(\zz-\yy))\sigma(\yy)\ \dif s_\yy\\
&= -\int_\Gamma {\bf t}_\xx\cdot\frac1\rho^2 G(\zz-\yy) {\bf t}_\yy \sigma(\yy)\ \dif s_\yy - 
({\bf t}_\xx\cdot\nabla_\zz)\int_\Gamma \frac{\partial G(\zz-\yy)}{\partial s_\yy}\sigma(\yy)\ \dif s_\yy\\
&= -\frac1\rho^2 {\bf t}_\xx\cdot \int_\Gamma G(\zz-\yy){\bf t}_\yy \sigma(\yy)\ \dif s_\yy + 
({\bf t}_\xx \cdot \nabla_\zz)\int_\Gamma G(\zz-\yy)\frac{\partial \sigma(\yy)}{\partial s}\ \dif s_\yy.
\end{align}
%
Here ${\bf t}_\xx$ is the tangent vector at the point $\xx$ and $d/ds$ is the arclength derivative.
Letting $\zz\to\xx\in\Gamma$, we obtain
%
\begin{equation}
\label{eq:normal_deriv}
\frac{\partial}{\partial \nnu_\xx} \DD[\sigma](\xx) = -\frac1\rho^2 {\bf t}_\xx\cdot \SSS[\sigma{\bf t}](\xx) + \frac{\partial}{\partial s}\SSS\left[\frac{\partial \sigma}{\partial s}\right](\xx).
\end{equation}
%

Substituting \eqref{eq:normal_deriv} to \eqref{eq:energy_BIE} provides the total energy calculation
throughout this paper and this expression takes the advantage of not calculating the quadruple layer potential.

@@

Distinct morphologies can be obtained from 
the HAP model by shifting the boundary condition $g(\xx)$.
Consider the linear case where $f(u) = u^2$.
This choice makes \eqref{eq:SL} a boundary value
problem for the screened-Laplace equation
and the solutions have a boundary layer structure that decays to $0$ in the
bulk with the decay length $\rho$.
We consider three kinds of boundary conditions
on particle $\Gamma_i$:
\begin{equation}
  \label{eq:bcs}
  \begin{tabular}{c|c|c}
     \text{(i)} & \text{(ii)} & \text{(iii)} \\
    \hline
    \rule{0pt}{4ex} 
      $g(\mathbf{x}) = \displaystyle \frac{1+\cos (\theta_i(\xx))}{\sqrt{3\pi r}}$
    & $g(\mathbf{x}) = \displaystyle \frac{2+\cos (\theta_i(\xx))}{\sqrt{9\pi r}}$
    & $g(\mathbf{x}) = \displaystyle \frac{\cos (\theta_i(\xx))}{\sqrt{\pi r}}$
\end{tabular}
\end{equation}
  The number $\theta_i(\xx)$ is the angle formed by the position $\xx$, the
particle center $\aa_i$, and the particle director $\dd_i$.% (Figure \ref{fig:flow_map}).
The normalizations e.g., $(\pi r)^{-1/2}$, provide a uniform
surface energy $\int_{\Gamma_i} g^2 \,ds = 1$ for circular particles
of radius $r$. 
The cases (ii) and (iii) are obtained by applying a vertical shift
and scaling to case (i), see Figure \ref{fig:bcs}.
Using this setup, we simulate the dynamics of 60 particles with one set of random
initial positions and orientations as shown in Figure~\ref{fig:relax}(a).

@@

\begin{figure}
  \begin{center}
  \includegraphics[width=0.4\textwidth]{Fig2a.png}
  \includegraphics[width=0.4\textwidth]{Fig2b.png}\\
  \includegraphics[width=0.4\textwidth]{Fig2c.png}
  \includegraphics[width=0.4\textwidth]{Fig2d.png}
  \end{center}
  \vspace{-20pt}  
  \caption{\label{fig:relax} (a) Initial configuration for $N_b=60$ structure. In order to expedite the self-assembly procedure, all particles are placed in a $8\times8$ box initially. (b)--(d) Equilibruim states for all boundary conditions~\eqref{eq:bcs}(i)--(iii), respectively. Colors represent the boundary values that increases from minimum (blue) to maximum (red).  Panel (b) reaches the equilibrium at $t=1200$; Panel (c) reaches the steady-state at $t=600$; Panel (d) reaches the final state at $t=180$.}
\end{figure}

Case (i) models amphiphilic particles.
Amphiphilic particles have a hydrophobic tail and a hydrophilic head
and these are accounted for as follows.
The hydrophilic side takes the value $u =0$.
This mimicks the apolar head of a lipid, for example, which does
not alter the structure of adjacent waters. The hydrophobic side
represents hydrocarbons and takes the value 
$u > 0$. The interaction between particles is attractive,
and particles will collectively orient their tails toward
one another. Figure~\ref{fig:relax}(b) demonstrates multiple bilayer structures such as pancake and vesicle shapes.

Multilamelar bilayers arise when both sides of the particle 
are hydrophobic. 
The boundary condition \eqref{eq:bcs}(ii) 
gives a particle with a hydrophobic intensity that is greater
on the $\theta_i = 0$ side than on the $\theta_i = \pi$ side.
The initial self-assembly is similar to that in case (i).
The difference arises in the long-time dynamics where the bilayers
no longer remain well-separated. 
Rather, the bilayers form layers 
as a consequence of the interfacial tension of exposed particle
heads.%~\cite{Huetal19, deMeetal21}. 
Figure~\ref{fig:relax}(c) shows the multilamellar structure with 8 layers when $N_b=60$. 
The number of folds depends on several factors, for instances, number of particles and intial configurations.

Finally, boundary condition \eqref{eq:bcs}(iii) models 
a particle whose head surface repels the tail surface as proposed in
\cite{MaRa76, Ma77}.
The particles initially form chains with their directors perpendicular to the
length of the chain. The equilibrium structure
resembles a checkerboard pattern
where each particle coordinates
its head with the head of three other particles and its tail with the
tail of three other particles. % Figure~\ref{fig:self-assembly}(c).
Figure~\ref{fig:relax}(d) gives an example that this boundary condition results a checker board/stripes equilibrium state.

@@

Consider placing equilibrium states of boundary condition \eqref{eq:bcs} (ii) for 60 particles in a shear flow,
\begin{equation}
\uu_\infty(\xx) = \dot\gamma ({\bf e}_y\cdot\xx){\bf e}_x,
\end{equation}
%
where $\dot\gamma$ is the shear rate. Figure~\ref{fig:shear_1} (a)--(c) show the snapshots of simulation results when $\dot\gamma=0.05$.
We measure the ratio of major and minor axes $\lambda$ by adopting an ellipse fitting using the singular value decomposition and track the ratio of this multi-lamellar structure when the flow is on. 
The result is shown in panel (d) and we observe that the multi-lamellar structure will be stretched 
to a pancake shape while the ratio reaches above $\lambda=3.5$ level.

Notice that for this simulation, the center of mass position of the initial configuration is not at the origin and it is slightly to the right of the origin. Therefore, the result has shown that the structure stretching occurs earlier that the case when we shift the center of mass position to the origin.

\begin{figure}
  \begin{center}
  \includegraphics[width=0.4\textwidth]{shear_0.png}
  \includegraphics[width=0.4\textwidth]{shear_1000.png}\\
  \includegraphics[width=0.4\textwidth]{shear_2000.png}
    \includegraphics[width=0.4\textwidth]{shear_dist_all.jpg}
  \end{center}
  \vspace{-20pt}  
  \caption{\label{fig:shear_1} Snapshots of simulation using boundary condition \eqref{eq:bcs}(ii) in a shear flow when $\dot\gamma=0.05$. }
\end{figure}

The second set of simulation is putting the equilibrium state of the boundary condition \eqref{eq:bcs} (iii), the checker board, in a shear flow. After a set of empirical numerical investigations, we obtain a critical shear rate $\dot\gamma=0.15$ where the checker board deforms into stripes and small groups of JP. Figure~\ref{fig:shear_2}(a)--(d) show configurations when $t=\{0,50,100,150\}$. %The deformation starts with stripe separations

@@
